---
title: "Introduction to chronam"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
library(chronam)
```

The goal of `chronam` is to make it easier to interface with Chronicling America's web API in R.

`chronam` provides built-in access to the metadata associated with Chronicling America as well as code for downloading the content of the digitized newspapers from the site.


## Chronicling America Metadata

Chronicling America's collection is stored in a hierarchical nested structure that can be explored by traversing many json files on the site.

Basically it can be represented as Batches > Titles (Newspapers) > Issues (Which contain pages, and within the pages ocr text blocks)

This package provides built-in versions of the metadata for Batches, Newspapers and Issues, as well as functions for re-downloading the metadata
from the Chronicling America site (if you suspect the package data is out of date).

### Batches

The list of batches can be accessed from the built-in package data,

```{r batches}
chronam::batches
```

If you want to ensure that the batches are completely up to date, they can be parsed from the chronicling america site by calling `get_batches()`
This function call can take several minutes to complete.

```{r get_batches, eval=FALSE}
batches = chronam::get_batches()
```

Source: https://chroniclingamerica.loc.gov/batches.json

### Titles (also refered to as Newspapers)

As with batches, the full list of newspapers can be accessed from the built-in package data.
```{r newpapers}
chronam::newspapers
```

Or can be generated with `get_newspapers()`
```{r get_newspapers}
newspapers = chronam::get_newspapers()
```

Source: https://chroniclingamerica.loc.gov/newspapers.json

### Issues

There are currently >2.6 million unique issues on Chronicling America. Note that the 'state' and 'year' column are added by this package and not in the original data. They are derived from the provided data and meant for convenience.

Load in the issues built-in package data.
```{r issues}
chronam::issues
```

It can take upwards of 30 minutes to download the full set from the site.
Regenerate the issues data:
```{r get_issues_all, eval=FALSE}
issues = chronam::get_issues_all(update=TRUE) #if update=FALSE will just use the package data
```

To get the issues for a specific lccn
```{r get_issues} 
lccn = "sn86072192"
issues = get_issues(lccn)
issues
```

Source: https://chroniclingamerica.loc.gov/lccn/<sn>. E.g https://chroniclingamerica.loc.gov/lccn/sn86072192.json

## Downloading Newspaper Content